TacticToe: Learning to Prove with Tactics

### 2018-02-20 tactictoe_jv.tex
Reread 8.2

We optimize \tactictoe by tuning parameters of six different techniques:
the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS 
evaluation and ATP integration. During training, optional techniques such as 
orthogonalization may also be turned off completely.
\tactictoe also includes other important techniques which are run with their 
default parameters: features extraction, prediction 
algorithms, number of premises for ATPs and MCTS evaluation radius (set to 
10).
By choosing a set of parameters, we create a strategy for 
\tactictoe that is evaluated on our training set of 273 theorems. And the 
strategy with the highest number of proven theorems is selected for 
a full scale evaluation.

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be raised from the 0.02 seconds used in \tactictoe's first
experiment~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for short
period of time. It seems that the alloted time allows a strategy
running without orthogonalization to catch up.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that have the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
on one thread asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
is given less predictions. But even if a lemma essential for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

The results shows that \tactictoe is able to prove almost twice as much
theorems as \eprover. Combining the results of \tactictoe and \eprover we get a
69.0\% success rate which is significantly above the 5O\% success rates
of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is
running a single set of parameters (strategy), where as hammers and ATPs have
been optimized and rely on a wide range of strategies.

\paragraph{Reconstruction}
Tactic proofs produced by \tactictoe during this experiment are all verifiable
in \holfour. By the design of the proof search, reconstruction of \tactictoe
proof succeeds, unless one of the tactic modifies the state of \holfour in a
way that changes the behavior of a tactic used in the final proof. Such an
event has not yet occurred.
More details on how a proof is extracted from the search tree is given in
Section~\ref{sec:proofdisplay}.
Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
by calling \metis for 2.0 seconds with the set of theorems used
in \eprover's proof as argument.

>>>>

We optimize \tactictoe by tuning parameters of six different techniques:
the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS 
evaluation and ATP integration. During training, optional techniques such as 
orthogonalization may also be turned off completely.
\tactictoe also includes other important techniques which are run with their 
default parameters: feature extraction, prediction 
algorithms, number of premises for ATPs and MCTS evaluation radius (set by default to
10).
By choosing a set of parameters, we create a strategy for 
\tactictoe that is evaluated on our training set of 273 theorems. And the 
strategy with the highest number of proven theorems is selected for 
a full scale evaluation.

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
experiments~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for short
period of time. It seems that the alloted time allows a strategy
running without orthogonalization to catch up.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that have the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
using a separate proces asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
is given less predictions. But even if an essential lemma for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

The results shows that \tactictoe is able to prove almost twice as many
theorems as \eprover. Combining the results of \tactictoe and \eprover we get a
69.0\% success rate which is significantly above the 50\% success rates
of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is
running a single set of parameters (strategy), where as hammers and ATPs have
been optimized and rely on a wide range of strategies.

\paragraph{Reconstruction}
Tactic proofs produced by \tactictoe during this experiment are all verifiable
in \holfour. By the design of the proof search, reconstruction of \tactictoe
proof succeeds, unless one of the tactic modifies the state of \holfour in a
way that changes the behavior of a tactic used in the final proof. This has not
occurred a single time in our experiments.
More details on how a proof is extracted from the search tree is given in
Section~\ref{sec:proofdisplay}.
Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
by calling \metis for 2.0 seconds with the set of theorems used
in \eprover's proof as argument.
